---
title: "Rexercise_week4"
author: Noémie
format: html
toc: True
---
#Exercise A Segmentation 

## Import data 
```{r}
library(readr)
wild<- read_delim("wildschwein_BE_2056.csv", delim=",")
```

## Task 1 Calculate distance
In the dataset Wildschwein, the trajectory of the wildschwein Rosa is going to be analysed. 
```{r}
 
library("sf")
library("dplyr")

rosa <- wild %>% 
    st_as_sf(coords = c("E", "N"), crs = 2056, remove = FALSE) %>% 
    filter(
      TierName == "Rosa", 
      DatetimeUTC >= "2014-11-07", 
      DatetimeUTC < "2014-11-17"
      )
```

### Step a)
Temporal window of 1.30 hours
--> including 6 fixes 
pos[n-3] to pos[n]
pos[n-2] to pos[n]
pos[n-1] to pos[n]
pos[n] to pos[n+1]
pos[n] to pos[n+2]
pos[n] to pos[n+3]
### Step b) 
Measure the distance to every point within v 
--> again with the function that we used for the demo 

```{r}
distance_by_element <- function(later, now) {
  as.numeric(
    st_distance(later, now, by_element = TRUE)
  )
}


rosa <- rosa  %>% 
    mutate(
        nMinus3 = distance_by_element(lag(geometry, n = 3), geometry),
        nMinus2 = distance_by_element(lag(geometry, n = 2), geometry),  # distance to pos -30 minutes --> two postitions before 
        nMinus1 = distance_by_element(lag(geometry, n = 1), geometry),  # distance to pos -15 minutes
        nPlus1  = distance_by_element(geometry, lead(geometry, 1)), # distance to pos +15 mintues
        nPlus2  = distance_by_element(geometry, lead(geometry, 2)),  # distance to pos +30 minutes
        nPlus3  = distance_by_element(geometry, lead(geometry, 3))
    )
        

rosa <- rosa %>%
    rowwise() %>% #one value per row
    mutate(
        stepMean = mean(c(nMinus3, nMinus2, nMinus1, nPlus1, nPlus2, nPlus3))
    ) %>%
    ungroup() #because of the rowwise! 

rosa
```
## Task 2 Specify and apply threshold d
visualize the stepmeans
```{r}
summary(rosa$stepMean)
boxplot(rosa$stepMean)
hist(rosa$stepMean)

rosa <- rosa %>% 
    mutate(static = stepMean < mean(stepMean, na.rm = TRUE))

```
## Task 3 Visualization 

```{r}
library(ggplot2)
ggplot(rosa)+
  geom_path(aes(E,N))+
  geom_sf(aes(color =static))
rosa_filter<- rosa %>% 
  filter(!static)
ggplot(rosa_filter)+
  geom_path(aes(E,N))+
  geom_sf(aes(color =static))
rosa_filter %>% 
    ggplot(aes(E, N)) +
    geom_point(data = rosa, col = "red") +
    geom_path() +
    geom_point() +
    coord_fixed() +
    theme(legend.position = "bottom")
```

## Task 4 Segment based analysis
Add a segment ID
```{r}
rle_id <- function(vec) {
    x <- rle(vec)$lengths
    as.factor(rep(seq_along(x), times = x))
}

rosa <- rosa %>% 
  mutate(segment_id = rle_id(static))
```

# Exercise B: Similarites
## Task 1: Similarities measures
import data 
```{r}
my_colors <- c("1" = "#D55E00", "2" = "#E69F00", "3" = "#009E73", 
               "4" = "#0072B2", "5" = "#56B4E9", "6" = "#CC79A7")
ped<-read_delim("pedestrian.csv")
ped
ped <- ped %>% 
    st_as_sf(coords = c("E", "N"), crs = 2056, remove = FALSE)
#visualization
ggplot(ped) +
  # 1. Couche de fond : TOUS les points en gris clair, pour TOUS les panneaux
  geom_sf(aes(group = TrajID), color = "grey85", size = 1.5) +
  
  # 2. Couche de premier plan : On trace TOUTES les trajectoires, 
  # mais on va les colorier selon leur ID.
  # Comme on est dans facet_wrap(~TrajID), ggplot va automatiquement 
  # "zoomer" sur l'ID correspondant, mais les données colorées sont toutes là.
  geom_sf(aes(color = factor(TrajID)), size = 2.5) +
  
  facet_wrap(~TrajID) +
  
  # On applique nos couleurs manuelles. 
  # Dans chaque panneau, seule la couleur correspondante sera visible au premier plan.
  scale_color_manual(values = my_colors, guide = "none") +
  
  labs(
    title = "Visual comparison of the 6 trajectories",
    subtitle = "Each subplot highlights a trajectory"
  ) +
  theme_minimal()
```
## Task 2: calculate similarity

The trajectories that are the most similar to me are: 
Traj 1 and 2; 6 and 1
```{r}
ped_no_geom <- st_drop_geometry(ped)
library(SimilarityMeasures)
#séparer les trajectoires en 6 dataset
traj_list <- split(ped_no_geom, ped_no_geom$TrajID)
#créer une matrixe avec chaque trajectoires 
traj_matrices <- lapply(traj_list, function(df) {
  coords <- df[, c("E", "N")]   # IMPORTANT: uniquement E et N
  coords$E <- as.numeric(coords$E)
  coords$N <- as.numeric(coords$N)
  as.matrix(coords)
})
#trier les points dans l'ordre
traj_list <- lapply(traj_list, function(df) {
  df[order(df$DatetimeUTC), ]
})
#check 
traj_matrices[[1]]
```
funtion to run all the 3 different methods of trajectory comparison between the traj 1 and the 5 others 
```{r}
results <- data.frame(
  Trajectory = paste0("traj", 2:6),
  DTW = sapply(2:6, function(i) DTW(traj_matrices[[1]], traj_matrices[[i]])),
  Frechet = sapply(2:6, function(i) Frechet(traj_matrices[[1]], traj_matrices[[i]])),
  EditDist = sapply(2:6, function(i) EditDist(traj_matrices[[1]], traj_matrices[[i]]))
)

results

```


Visualisation 
```{r}
#| label: fig-plot-result
#| fig-cap: "comparison of similarity measures"

library(ggplot2)
library(tidyr)

results_long <- pivot_longer(results, 
                             cols = -Trajectory,
                             names_to = "Measure",
                             values_to = "Distance")

ggplot(results_long, aes(x = Trajectory, y = Distance, fill = Measure)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  ggtitle("Trajectory Similarity Comparison (without normalization of the data")
```
Visualisation avec des données log 
```{r}
#| label: fig-plot-result-log
#| fig-cap: "comparison of similarity measures (log-scale)"
results_long <- pivot_longer(results, 
                             cols = -Trajectory,
                             names_to = "Measure",
                             values_to = "Distance")

ggplot(results_long, aes(x = Trajectory, y = Distance, fill = Measure)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_y_log10() +
  theme_minimal() +
  ggtitle("Trajectory Similarity (log scale)")
```

### Results
With the plots @fig-plot-result and @fig-plot-result-log, it is possible to see that the trajectory 6 is the most similar to the trajectory 1 and the trajectory 3 is the most dissimilar. The logaritmization of the values was necessary since the range of values are very different depending on the test executed. 

