---
title: "Rexercise_week4"
author: Noémie
format: html
toc: True
---
#Exercise A Segmentation 

## Import data 
```{r}
library(readr)
wild<- read_delim("wildschwein_BE_2056.csv", delim=",")
```

## Task 1 Calculate distance
In the dataset Wildschwein, the trajectory of the wildschwein Rosa is going to be analysed. 
```{r}
 
library("sf")
library("dplyr")

rosa <- wild %>% 
    st_as_sf(coords = c("E", "N"), crs = 2056, remove = FALSE) %>% 
    filter(
      TierName == "Rosa", 
      DatetimeUTC >= "2014-11-07", 
      DatetimeUTC < "2014-11-17"
      )
```

### Step a)
Temporal window of 1.30 hours
--> including 6 fixes 
pos[n-3] to pos[n]
pos[n-2] to pos[n]
pos[n-1] to pos[n]
pos[n] to pos[n+1]
pos[n] to pos[n+2]
pos[n] to pos[n+3]
### Step b) 
Measure the distance to every point within v 
--> again with the function that we used for the demo 

```{r}
distance_by_element <- function(later, now) {
  as.numeric(
    st_distance(later, now, by_element = TRUE)
  )
}


rosa <- rosa  %>% 
    mutate(
        nMinus3 = distance_by_element(lag(geometry, n = 3), geometry),
        nMinus2 = distance_by_element(lag(geometry, n = 2), geometry),  # distance to pos -30 minutes --> two postitions before 
        nMinus1 = distance_by_element(lag(geometry, n = 1), geometry),  # distance to pos -15 minutes
        nPlus1  = distance_by_element(geometry, lead(geometry, 1)), # distance to pos +15 mintues
        nPlus2  = distance_by_element(geometry, lead(geometry, 2)),  # distance to pos +30 minutes
        nPlus3  = distance_by_element(geometry, lead(geometry, 3))
    )
        

rosa <- rosa %>%
    rowwise() %>% #one value per row
    mutate(
        stepMean = mean(c(nMinus3, nMinus2, nMinus1, nPlus1, nPlus2, nPlus3))
    ) %>%
    ungroup() #because of the rowwise! 

rosa
```
## Task 2 Specify and apply threshold d
visualize the stepmeans
```{r}
summary(rosa$stepMean)
boxplot(rosa$stepMean)
hist(rosa$stepMean)

rosa <- rosa %>% 
    mutate(static = stepMean < mean(stepMean, na.rm = TRUE))

```
## Task 3 Visualization 

```{r}
library(ggplot2)
ggplot(rosa)+
  geom_path(aes(E,N))+
  geom_sf(aes(color =static))
rosa_filter<- rosa %>% 
  filter(!static)
ggplot(rosa_filter)+
  geom_path(aes(E,N))+
  geom_sf(aes(color =static))
rosa_filter %>% 
    ggplot(aes(E, N)) +
    geom_point(data = rosa, col = "red") +
    geom_path() +
    geom_point() +
    coord_fixed() +
    theme(legend.position = "bottom")
```

## Task 4 Segment based analysis
Add a segment ID
```{r}
rle_id <- function(vec) {
    x <- rle(vec)$lengths
    as.factor(rep(seq_along(x), times = x))
}

rosa <- rosa %>% 
  mutate(segment_id = rle_id(static))
```

# Exercise B: Similarites
## Task 1: Similarities measures
import data 
```{r}
ped<-read_delim("pedestrian.csv")

```

